---
title: "Delta Smelt Occurrence by Life Stage"
author: "Brian Mahardja"
date: "10/04/2022"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = 'docs',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#devtools::install_github("sbashevkin/LTMRdata")

library(LTMRdata)
library(tidyverse)
library(lubridate)
```

## Load and process Delta Smelt data

Grab Delta Smelt data from the integrated fish dataset (LTMRdata package by Sam Bashevkin et al.) and produce one fish data per row. I will exclude fish with no length measurement. Note that unmeasured fish that were collected alongside measured fish were converted to have fish length per LTMRdata documentation. I converted count to integer however, since the process that Sam implemented created decimal points for counts.

Surveys included in this calculation:

1. San Francisco Bay Study

2. Suisun Marsh Study

3. Fall Midwater Trawl

4. Spring Kodiak Trawl

5. Delta Juvenile Fish Monitoring Program

6. Enhanced Delta Smelt Monitoring

7. 20-mm Survey

8. Smelt Larval Survey

9. Summer Townet Survey

```{r data, echo=TRUE, message=FALSE, warning=FALSE}

#Load data
data_deltasmelt <-LTMRdata::fish(sources=c("Baystudy", "Suisun", "FMWT", "SKT", "DJFMP", "EDSM", "TMM", "SLS","STN"),species="Hypomesus transpacificus", size_cutoff=NULL,remove_unknown_lengths=TRUE) %>% filter(year(Date)>=1994)

#Create time series summary
data_deltasmelt_time_summary<-data_deltasmelt %>% group_by(Source) %>% summarise(StartDate=min(Date),EndDate=max(Date))
data_deltasmelt_time_summary

#Change count to integer
data_deltasmelt$Count<-as.integer(data_deltasmelt$Count)
#Remove zero catch data
data_deltasmelt<- data_deltasmelt%>% filter(Count>0)

#Create long form, remove data prior to 1994
data_deltasmelt_long <-data_deltasmelt[rep(1:nrow(data_deltasmelt), data_deltasmelt$Count),] %>% mutate(Count=1)

```

## Size cutoffs used for Delta Smelt

Larvae: (<20mm FL)

Juvenile: (20-58mm FL)

Adult: (>58mm FL)

### Brood year cutoff for the life stages

Larvae: brood year = calendar year

Juvenile: brood year starts in March 1st of current year to February 28th or 29th of the following year

Adult: brood year starts in June 1st of current year to May 31st of the following year

## Adult Delta Smelt 

(>58 mm FL, cohort year set to June 1st - May 31st)


```{r, echo=FALSE, message=FALSE, warning=FALSE}

col_order <- c("broodyear", "0.0%", "5.0%",
               "10.0%", "90.0%","95.0%","100.0%")

data_deltasmelt_adult<-data_deltasmelt_long %>% filter(Length>58) %>% mutate(broodyear=ifelse(month(Date)>=6,year(Date),year(Date)-1))

data_deltasmelt_adult_sum<-data_deltasmelt_adult %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_deltasmelt_adult_sum$date<-format(data_deltasmelt_adult_sum$date, format="%B %d %Y")

data_deltasmelt_adult_sum<-spread(data_deltasmelt_adult_sum,quantile,date)

data_deltasmelt_adult_sum <- data_deltasmelt_adult_sum[, col_order]

data_deltasmelt_adult_sum <- left_join(data_deltasmelt_adult_sum,data_deltasmelt_adult %>% group_by(broodyear) %>% summarise(sample_size=sum(Count))) %>% filter(broodyear>=1994)

knitr::kable(data_deltasmelt_adult_sum,col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

Note that adult Delta Smelt can linger for more than a year (i.e, 1+ year old Delta Smelt). 2-year old Delta Smelt have even been observed in the past. This is why the first adult smelt can sometimes show up in June or July in the table. This phenomenon of 1+ year old Delta Smelt appear to be more common in the 1990s (or whenever smelt were more abundant). Example below:

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data_deltasmelt_long %>% filter(year(Date)==1994), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="1994 Delta Smelt FL Distribution",x="Date")

ggplot(data_deltasmelt_long %>% filter(year(Date)==1998), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="1998 Delta Smelt FL Distribution",x="Date")

```


## Juvenile Delta Smelt 

(20-58 mm FL, cohort year set to March 1st - February 28th/29th)


```{r, echo=FALSE, message=FALSE, warning=FALSE}

data_deltasmelt_juvenile<-data_deltasmelt_long %>% filter(Length<=58&Length>=20) %>% mutate(broodyear=ifelse(month(Date)>=3,year(Date),year(Date)-1))

data_deltasmelt_juvenile_sum<-data_deltasmelt_juvenile %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_deltasmelt_juvenile_sum$date<-format(data_deltasmelt_juvenile_sum$date, format="%B %d %Y")

data_deltasmelt_juvenile_sum<-spread(data_deltasmelt_juvenile_sum,quantile,date)

data_deltasmelt_juvenile_sum <- data_deltasmelt_juvenile_sum[, col_order]

data_deltasmelt_juvenile_sum <- left_join(data_deltasmelt_juvenile_sum,data_deltasmelt_juvenile %>% group_by(broodyear) %>% summarise(sample_size=sum(Count)))%>% filter(broodyear>=1994)

knitr::kable(data_deltasmelt_juvenile_sum,col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

## Larval Delta Smelt 

(<20 mm FL, cohort year set to just calendar year)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

data_deltasmelt_larvae<-data_deltasmelt_long %>% filter(Length<=20) %>% mutate(broodyear=year(Date))

data_deltasmelt_larvae_sum<-data_deltasmelt_larvae %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_deltasmelt_larvae_sum$date<-format(data_deltasmelt_larvae_sum$date, format="%B %d %Y")

data_deltasmelt_larvae_sum<-spread(data_deltasmelt_larvae_sum,quantile,date)

data_deltasmelt_larvae_sum <- data_deltasmelt_larvae_sum[, col_order]

data_deltasmelt_larvae_sum <- left_join(data_deltasmelt_larvae_sum,data_deltasmelt_larvae %>% group_by(broodyear) %>% summarise(sample_size=sum(Count)))

knitr::kable(data_deltasmelt_larvae_sum,col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

