---
title: "Delta Smelt Occurrence by Life Stage"
author: "Brian Mahardja"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#devtools::install_github("sbashevkin/LTMRdata")

library(LTMRdata)
library(tidyverse)
library(lubridate)
```

## Load and process Delta Smelt data

Grab Delta Smelt data from the integrated fish dataset (LTMRdata package by Sam Bashevkin et al.) and produce one fish data per row. I will exclude fish with no length measurement. Note that unmeasured fish that were collected alongside measured fish were converted to have fish length per LTMRdata documentation. I converted count to integer however, since the process that Sam implemented created decimal points for counts.

Surveys included in this calculation:

1. San Francisco Bay Study

2. Suisun Marsh Study

3. Fall Midwater Trawl

4. Spring Kodiak Trawl

5. Delta Juvenile Fish Monitoring Program

6. Enhanced Delta Smelt Monitoring

7. 20-mm Survey

8. Smelt Larval Survey

9. Summer Townet Survey

```{r data, echo=TRUE, message=FALSE, warning=FALSE}

#Load data
data_deltasmelt <-LTMRdata::fish(sources=c("Baystudy", "Suisun", "FMWT", "SKT", "DJFMP", "EDSM", "TMM", "SLS","STN"),species="Hypomesus transpacificus", size_cutoff=NULL,remove_unknown_lengths=TRUE) %>%
  filter(Count>0)

#Change count to integer
data_deltasmelt$Count<-as.integer(data_deltasmelt$Count)

#Create long form, remove data prior to 1994
data_deltasmelt_long <-data_deltasmelt[rep(1:nrow(data_deltasmelt), data_deltasmelt$Count),] %>% mutate(Count=1) %>% filter(year(Date)>=1994)


```

## Size cutoffs used for Delta Smelt

Larvae: (<20mm FL)

Juvenile: (20-58mm FL)

Adult: (>58mm FL)

### Brood year cutoff for the life stages

Larvae: brood year = calendar year

Juvenile: brood year starts in March 1st of current year to February 28th or 29th of the following year

Adult: brood year starts in June 1st of current year to May 31st of the following year

```{r}

col_order <- c("broodyear", "0.0%", "5.0%",
               "10.0%", "90.0%","95.0%","100.0%")

data_deltasmelt_adult<-data_deltasmelt_long %>% filter(Length>=55) %>% mutate(broodyear=ifelse(month(Date)>=6,year(Date),year(Date)-1))

data_deltasmelt_adult_sum<-data_deltasmelt_adult %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_deltasmelt_adult_sum$date<-format(data_deltasmelt_adult_sum$date, format="%B %d %Y")

data_deltasmelt_adult_sum<-spread(data_deltasmelt_adult_sum,quantile,date)

data_deltasmelt_adult_sum <- data_deltasmelt_sum[, col_order]

data_deltasmelt_adult_sum <- left_join(data_deltasmelt_adult_sum,data_deltasmelt_adult %>% group_by(broodyear) %>% summarise(sample_size=sum(Count)))


data_deltasmelt_adult_sum
```

