---
title: "Longfin Smelt Occurrence by Life Stage"
author: "Brian Mahardja"
date: "11/19/2021"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = 'docs',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#devtools::install_github("sbashevkin/LTMRdata")

library(LTMRdata)
library(tidyverse)
library(lubridate)
```


## Load and process Longfin Smelt data

Grab Longfin Smelt data from the integrated fish dataset (LTMRdata package by Sam Bashevkin et al.) and produce one fish data per row. I will exclude fish with no length measurement. Note that unmeasured fish that were collected alongside measured fish were converted to have fish length per LTMRdata documentation. I converted count to integer however, since the process that Sam implemented created decimal points for counts.

Surveys included in this calculation:

1. San Francisco Bay Study

2. Suisun Marsh Study

3. Fall Midwater Trawl

4. Spring Kodiak Trawl

5. Delta Juvenile Fish Monitoring Program

6. Enhanced Delta Smelt Monitoring

7. 20-mm Survey

8. Smelt Larval Survey

9. Summer Townet Survey

```{r data, echo=TRUE, message=FALSE, warning=FALSE}

#Load data
data_longfinsmelt <-LTMRdata::fish(sources=c("Baystudy", "Suisun", "FMWT", "SKT", "DJFMP", "EDSM", "TMM", "SLS","STN"),species="Spirinchus thaleichthys", size_cutoff=NULL,remove_unknown_lengths=TRUE) %>%
  filter(Count>0)

#Change count to integer
data_longfinsmelt$Count<-as.integer(data_longfinsmelt$Count)

#Create long form, remove data prior to 1994
data_longfinsmelt_long <-data_longfinsmelt[rep(1:nrow(data_longfinsmelt), data_longfinsmelt$Count),] %>% mutate(Count=1) %>% filter(year(Date)>=1994)

```


## Size cutoffs used for Longfin Smelt

Larvae: (<20mm FL)

Juvenile: (20-84mm FL)

Adult: (>84mm FL)

### Brood year cutoff for the life stages

Larvae: brood year = calendar year

Juvenile: brood year = calendar year

Adult: brood year starts in July 1st of current year to June 30th of the following year

# Overview of size distribution

Subadult and adult Longfin Smelt typically are present and caught from January to July, and then again starting October to November.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data_longfinsmelt_long %>% filter(year(Date)==2017), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="2017 Longfin Smelt FL Distribution",x="Date")

ggplot(data_longfinsmelt_long %>% filter(year(Date)==2018), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="2018 Longfin Smelt FL Distribution",x="Date")

ggplot(data_longfinsmelt_long %>% filter(year(Date)==2019), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="2019 Longfin Smelt FL Distribution",x="Date")


```

## Adult Longfin Smelt 

(mature and immature adults of >84 mm FL, cohort year set to July 1st - June 30th)


```{r, echo=FALSE, message=FALSE, warning=FALSE}

col_order <- c("broodyear", "0.0%", "5.0%",
               "10.0%", "90.0%","95.0%","100.0%")

data_longfinsmelt_adult<-data_longfinsmelt_long %>% filter(Length>84) %>% mutate(broodyear=ifelse(month(Date)>=7,year(Date),year(Date)-1))

data_longfinsmelt_adult_sum<-data_longfinsmelt_adult %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_longfinsmelt_adult_sum$date<-format(data_longfinsmelt_adult_sum$date, format="%B %d %Y")

data_longfinsmelt_adult_sum<-spread(data_longfinsmelt_adult_sum,quantile,date)

data_longfinsmelt_adult_sum <- data_longfinsmelt_adult_sum[, col_order]

data_longfinsmelt_adult_sum <- left_join(data_longfinsmelt_adult_sum,data_longfinsmelt_adult %>% group_by(broodyear) %>% summarise(sample_size=sum(Count))) %>% filter(broodyear>=1994&broodyear<2021)

knitr::kable(data_longfinsmelt_adult_sum,col.names = c("'Cohort' Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

## Juvenile Longfin Smelt 

(20-84 mm FL, cohort year set to January 1st - December 31st; same as calendar year)


```{r, echo=FALSE, message=FALSE, warning=FALSE}

data_longfinsmelt_juvenile<-data_longfinsmelt_long %>% filter(Length<=84&Length>=20) %>% mutate(broodyear=year(Date))

data_longfinsmelt_juvenile_sum<-data_longfinsmelt_juvenile %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_longfinsmelt_juvenile_sum$date<-format(data_longfinsmelt_juvenile_sum$date, format="%B %d %Y")

data_longfinsmelt_juvenile_sum<-spread(data_longfinsmelt_juvenile_sum,quantile,date)

data_longfinsmelt_juvenile_sum <- data_longfinsmelt_juvenile_sum[, col_order]

data_longfinsmelt_juvenile_sum <- left_join(data_longfinsmelt_juvenile_sum,data_longfinsmelt_juvenile %>% group_by(broodyear) %>% summarise(sample_size=sum(Count)))%>% filter(broodyear>=1994)

knitr::kable(data_longfinsmelt_juvenile_sum,col.names = c("Calendar Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```


## Larval Longfin Smelt 

(<20 mm FL, cohort year set to just calendar year)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

data_longfinsmelt_larvae<-data_longfinsmelt_long %>% filter(Length<=20) %>% mutate(broodyear=year(Date))

data_longfinsmelt_larvae_sum<-data_longfinsmelt_larvae %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_longfinsmelt_larvae_sum$date<-format(data_longfinsmelt_larvae_sum$date, format="%B %d %Y")

data_longfinsmelt_larvae_sum<-spread(data_longfinsmelt_larvae_sum,quantile,date)

data_longfinsmelt_larvae_sum <- data_longfinsmelt_larvae_sum[, col_order]

data_longfinsmelt_larvae_sum <- left_join(data_longfinsmelt_larvae_sum,data_longfinsmelt_larvae %>% group_by(broodyear) %>% summarise(sample_size=sum(Count)))

knitr::kable(data_longfinsmelt_larvae_sum,col.names = c("Calendar Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

## Adult Longfin Smelt Migration into the Delta

Per conversation with Trinh Nguyen, who is working on a Longfin Smelt migration window model using Chipps Island Trawl data with Jim Hobbs and Michael Eakin, 2+ year old Longfin Smelt catch are defined as "spawning adults" moving into the Delta.

Size cut-off for their study was >=98 mm FL. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}

data_longfinsmelt_migration<-data_longfinsmelt_long %>% filter(Length>=98, Source=="DJFMP", grepl("SB", Station), Method=="Midwater trawl") %>% mutate(broodyear=ifelse(month(Date)<10,year(Date),year(Date)+1))


data_longfinsmelt_migration_sum<-data_longfinsmelt_migration %>% group_by(broodyear) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_longfinsmelt_migration_sum$date<-format(data_longfinsmelt_migration_sum$date, format="%B %d %Y")

data_longfinsmelt_migration_sum<-spread(data_longfinsmelt_migration_sum,quantile,date)

data_longfinsmelt_migration_sum <- data_longfinsmelt_migration_sum[, col_order]

data_longfinsmelt_migration_sum <- left_join(data_longfinsmelt_migration_sum,data_longfinsmelt_migration %>% group_by(broodyear) %>% summarise(sample_size=sum(Count))) %>% filter(broodyear>1994)

knitr::kable(data_longfinsmelt_migration_sum,col.names = c("Water Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```
