---
title: "Longfin Smelt Occurrence by Life Stage"
author: "Brian Mahardja"
date: "`r Sys.Date()`"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = 'docs',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#devtools::install_github("sbashevkin/LTMRdata")
#devtools::install_github("InteragencyEcologicalProgram/deltamapr")

library(deltafish)
library(deltamapr)
library(tidyverse)
library(lubridate)
library(sf)
```


## Load and process Longfin Smelt data

Grab Longfin Smelt data from the integrated fish dataset (LTMRdata package by Sam Bashevkin et al.) and produce one fish data per row. I will exclude fish with no length measurement. Note that unmeasured fish that were collected alongside measured fish were converted to have fish length per LTMRdata documentation. I converted count to integer however, since the process that Sam implemented created decimal points for counts.

Surveys included in this calculation:

1. San Francisco Bay Study

2. Suisun Marsh Study

3. Fall Midwater Trawl

4. Spring Kodiak Trawl

5. Delta Juvenile Fish Monitoring Program

6. Enhanced Delta Smelt Monitoring

7. 20-mm Survey

8. Smelt Larval Survey

9. Summer Townet Survey

```{r data, echo=TRUE, message=FALSE, warning=FALSE}


#Use deltafish to load data
surv <- open_survey()
fish <- open_fish()

# filter for taxa of interest
fish_longfinsmelt <- fish %>% 
  filter(Taxa %in% c("Spirinchus thaleichthys"))


# do a join and collect the resulting data frame
# collect executes the sql query and gives you a table
data_longfinsmelt<-left_join(surv, fish_longfinsmelt) %>% 
  collect() 


#Create time series summary
data_longfinsmelt_time_summary<-data_longfinsmelt %>% group_by(Source) %>% summarise(StartDate=min(Date),EndDate=max(Date))
data_longfinsmelt_time_summary


#Remove 0 counts, only use years 1994 and after, only ones with non-zero lengths
data_longfinsmelt <- data_longfinsmelt %>% 
  filter(Count>0) %>% filter(year(Date)>=1994) %>% filter(!is.na(Length))

#Change count to integer
data_longfinsmelt$Count<-as.integer(data_longfinsmelt$Count)
#Remove zero catch data
data_longfinsmelt<- data_longfinsmelt%>% filter(Count>0)

#Create long form, remove data prior to 1994
data_longfinsmelt_long <-data_longfinsmelt[rep(1:nrow(data_longfinsmelt), data_longfinsmelt$Count),] %>% mutate(Count=1)
```
## Regions used to understand spatial distribution

Shown below are the regional cutoffs used to better understand when fish are found in a particular area within the estuary

```{r map, echo=FALSE, message=FALSE, warning=FALSE}
subregion_def<-read_csv(file="subregion_to_stratum_17_1.csv")
highrisk<- subregion_def %>% filter(grepl("HR",Stratum)) %>% select(SubRegion)
  
highrisk

custom_region <-R_EDSM_Subregions_Mahardja %>% mutate(CustomRegion = case_when(
  SubRegion %in% c("South Bay","San Francisco Bay","San Pablo Bay") ~ "Bays",
  SubRegion %in% highrisk$SubRegion ~ "Central and South Delta",
  TRUE ~ "North Delta and Suisun Bay"))

ggplot()+
  geom_sf(data=WW_Delta,fill="grey")+
  geom_sf(data=custom_region,aes(color=CustomRegion,fill=CustomRegion),alpha=0.4)+
  theme_bw()+ theme(legend.title = element_blank())

```

## Create spatial data frame, categorizing fish catch into regions

```{r spatial, echo=FALSE, message=FALSE, warning=FALSE }

data_longfinsmelt_long_spatial<- data_longfinsmelt_long %>% filter(!is.na(Latitude)&!is.na(Longitude)) %>%
  st_as_sf(coords=c("Longitude", "Latitude"),crs=st_crs(WW_Delta))

data_longfinsmelt_long_spatial<- st_transform(data_longfinsmelt_long_spatial, crs = st_crs(custom_region))

data_longfinsmelt_long_spatial<- st_join(data_longfinsmelt_long_spatial,custom_region)

st_geometry(data_longfinsmelt_long_spatial) <- NULL # remove geometry, coerce to data.frame

```

## Size cutoffs used for Longfin Smelt

Larvae: (<20mm FL)

Juvenile: (20-84mm FL)

Adult: (>84mm FL)

### Brood year cutoff for the life stages

Larvae: brood year = calendar year

Juvenile: brood year = calendar year

Adult: brood year starts in July 1st of current year to June 30th of the following year

# Overview of size distribution

Subadult and adult Longfin Smelt typically are present and caught from January to July, and then again starting October to November.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data_longfinsmelt_long %>% filter(year(Date)==2017), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="2017 Longfin Smelt FL Distribution",x="Date")

ggplot(data_longfinsmelt_long %>% filter(year(Date)==2018), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="2018 Longfin Smelt FL Distribution",x="Date")

ggplot(data_longfinsmelt_long %>% filter(year(Date)==2019), aes(x=Date, y=Length)) + theme_bw() +
  geom_point(size=1, shape=16,alpha=0.2) + labs(title="2019 Longfin Smelt FL Distribution",x="Date")


```


## Adult Longfin Smelt 

(mature and immature adults of >84 mm FL, cohort year set to July 1st - June 30th)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

col_order <- c("broodyear","CustomRegion", "0.0%", "5.0%",
               "10.0%", "90.0%","95.0%","100.0%")

data_longfinsmelt_adult<-data_longfinsmelt_long_spatial %>% filter(Length>84) %>% mutate(broodyear=ifelse(month(Date)>=7,year(Date),year(Date)-1))

data_longfinsmelt_adult_sum<-data_longfinsmelt_adult %>% group_by(broodyear,CustomRegion) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_longfinsmelt_adult_sum$date<-format(data_longfinsmelt_adult_sum$date, format="%B %d %Y")

data_longfinsmelt_adult_sum<-spread(data_longfinsmelt_adult_sum,quantile,date)

data_longfinsmelt_adult_sum <- data_longfinsmelt_adult_sum[, col_order]

data_longfinsmelt_adult_sum <- left_join(data_longfinsmelt_adult_sum,data_longfinsmelt_adult %>% group_by(broodyear,CustomRegion) %>% summarise(sample_size=sum(Count))) %>% filter(broodyear>=1994)


```


## Juvenile Longfin Smelt 

(20-84 mm FL, cohort year set to January 1st - December 31st; same as calendar year)


```{r, echo=FALSE, message=FALSE, warning=FALSE}

data_longfinsmelt_juvenile<-data_longfinsmelt_long_spatial %>% filter(Length<=84&Length>=20) %>% mutate(broodyear=year(Date))


data_longfinsmelt_juvenile_sum<-data_longfinsmelt_juvenile %>% group_by(broodyear,CustomRegion) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_longfinsmelt_juvenile_sum$date<-format(data_longfinsmelt_juvenile_sum$date, format="%B %d %Y")

data_longfinsmelt_juvenile_sum<-spread(data_longfinsmelt_juvenile_sum,quantile,date)

data_longfinsmelt_juvenile_sum <- data_longfinsmelt_juvenile_sum[, col_order]

data_longfinsmelt_juvenile_sum <- left_join(data_longfinsmelt_juvenile_sum,data_longfinsmelt_juvenile %>% group_by(broodyear,CustomRegion) %>% summarise(sample_size=sum(Count))) %>% filter(broodyear>=1994)


```


## Larval Longfin Smelt 

(<20 mm FL, cohort year set to just calendar year)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

data_longfinsmelt_larvae<-data_longfinsmelt_long_spatial %>% filter(Length<=20) %>% mutate(broodyear=year(Date))

data_longfinsmelt_larvae_sum<-data_longfinsmelt_larvae %>% group_by(broodyear,CustomRegion) %>%  
  summarise(quantile = scales::percent(c(0, 0.05, 0.1, 0.9, 0.95, 1.0)),
            date = quantile(Date, c(0, 0.05, 0.1, 0.9, 0.95, 1.0), type=1)) 

data_longfinsmelt_larvae_sum$date<-format(data_longfinsmelt_larvae_sum$date, format="%B %d %Y")

data_longfinsmelt_larvae_sum<-spread(data_longfinsmelt_larvae_sum,quantile,date)

data_longfinsmelt_larvae_sum <- data_longfinsmelt_larvae_sum[, col_order]

data_longfinsmelt_larvae_sum <- left_join(data_longfinsmelt_larvae_sum,data_longfinsmelt_larvae %>% group_by(broodyear,CustomRegion) %>% summarise(sample_size=sum(Count))) %>% filter(broodyear>=1994)


```


## Bay Region
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot()+
  geom_sf(data=WW_Delta,fill="grey")+
  geom_sf(data=custom_region %>% filter(CustomRegion=="Bays"),color="blue", fill="blue",alpha=0.4)+
  theme_bw()+ theme(legend.title = element_blank())

```
### Adults

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_adult_sum %>% filter(CustomRegion=="Bays") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```
### Juveniles

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_juvenile_sum %>% filter(CustomRegion=="Bays") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```
### Larvae

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_larvae_sum %>% filter(CustomRegion=="Bays") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```



## North Delta and Suisun Bay Region
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot()+
  geom_sf(data=WW_Delta,fill="grey")+
  geom_sf(data=custom_region %>% filter(CustomRegion=="North Delta and Suisun Bay"),color="green", fill="green",alpha=0.4)+
  theme_bw()+ theme(legend.title = element_blank())

```

### Adults

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_adult_sum %>% filter(CustomRegion=="North Delta and Suisun Bay") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

### Juveniles

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_juvenile_sum %>% filter(CustomRegion=="North Delta and Suisun Bay") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

### Larvae

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_larvae_sum %>% filter(CustomRegion=="North Delta and Suisun Bay") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```

## Central and South Delta
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot()+
  geom_sf(data=WW_Delta,fill="grey")+
  geom_sf(data=custom_region %>% filter(CustomRegion=="Central and South Delta"),color="red", fill="green",alpha=0.4)+
  theme_bw()+ theme(legend.title = element_blank())

```

### Adults

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_adult_sum %>% filter(CustomRegion=="Central and South Delta") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```
### Juveniles

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_juvenile_sum %>% filter(CustomRegion=="Central and South Delta") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```
### Larvae

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(data_longfinsmelt_larvae_sum %>% filter(CustomRegion=="Central and South Delta") %>% ungroup() %>% select(-CustomRegion),col.names = c("Cohort Year","0.0%",
                                              "5.0%","10.0%", "90.0%",
                                              "95.0%","100.0%","Sample size"))

```
